---
title: "Ho Viet"
author: "Duc Du"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls(all.names = TRUE))
knitr::opts_chunk$set(echo = FALSE, 
                      cache=TRUE,
                      tidy.opts=list(width.cutoff=60),
                      tidy=TRUE)
#library(rmarkdown)
#library(knitr)
library(markdown)
library(stringr)
library(forcats)
library(RColorBrewer)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(Hmisc)
library(gtsummary)
#library(flextable)
library(huxtable)
library(data.table)
library(lubridate)
library(stringi)
library(forcats)
library(ggpubr)
library(rstatix)
#library(dlookr)
#library(flextable)
library(haven)
library(raster)
library(rgdal)
library(viridis)
library(sf)

select<-dplyr::select
filter<-dplyr::filter
recode<-dplyr::recode
getData<-raster::getData

# rotate function (see here: https://r-spatial.github.io/sf/articles/sf3.html#affine-transformations
rot <- function(a) matrix(c(cos(a), sin(a), -sin(a), cos(a)), 2, 2)

options(scipen=999, digits = 3)
options(max.print=1000000)
```

```{r}
# if Shiny is not installed, install it.
if (!require("shiny")) { install.packages("shiny", repos="http://cran.us.r-project.org") }
    
# Load the Shiny library
library(shiny)

# Set `leaflet` as our current working directory.
setwd("D:/Viet/Thong-ke-Ho-Viet/shiny/leaflet")
```

```{r}
# Now let’s create our starter files
file.create("global.R")
file.create("server.R")
file.create("ui.R")
```

# Load data

```{r load data2}
load("../../data/hoviet.RData")
```

# Creating our Map

We will be using Leaflet to create our map, so we will need to include the leaflet library. First let’s install the leaflet package.

```{r}
if (!require("leaflet")) { install.packages("leaflet", repos="http://cran.us.r-project.org")}
library(leaflet)
```

We will start with the global.R file. The global.R file stores all our global variables. Let’s edit it.

```{r}
file.edit("global.R")
```

Copy the following into the global.R file:

```{r}
# Loads the Shiny and leaflet libraries.
library(shiny)
library(leaflet)

# Saves the breakfast cereal dataset to the `cereals` variable.
crime <- read.csv("crime.csv", header=TRUE, stringsAsFactors=FALSE)
```

Save and close the global.R file. Let’s setup our server.R file. Open up the file.

```{r}
file.edit("server.R")
```


```{r}
library(leaflet)
mai_sff <- dat_sff %>% 
  mutate(prop.ho=(sum(subset(., Ho=="Mai")[, "songuoi"])/sum(songuoi))*100) %>% 
  filter(Ho=="Mai") %>% 
  st_as_sf(crs = 4326)

mai_sff$centroid <- sf::st_transform(mai_sff, 2154) %>% 
  st_centroid() %>% 
  # this is the crs from d, which has no EPSG code:
  st_transform(., '+proj=longlat +ellps=GRS80 +no_defs') %>%
  # since you want the centroids in a second geometry col:
  st_geometry()

mai <- mai_sff %>%
  mutate(x = map(centroid,1),
         y = map(centroid,2),
         x = as.numeric(x),
         y = as.numeric(y))

pal <- colorNumeric(palette = "viridis", reverse = TRUE, domain = mai_sff$songuoi_km, alpha = TRUE)

mai_sff %>%
    st_transform(crs = "+init=epsg:4326") %>%
    leaflet(width = "100%") %>%
    addProviderTiles(provider = "Esri.WorldStreetMap") %>%
    addPolygons(popup = ~ str_extract(NAME_3, "^([^,]*)"),
                stroke = FALSE,
                smoothFactor = 0,
                fillOpacity = 0.4,
                color = ~ pal(songuoi_km)) %>%
    addLegend("bottomright", 
              pal = pal, 
              values = ~ songuoi_km,
              title = "Mật độ người/km2",
              opacity = 1)

save(mai, file = "D:/Viet/Thong-ke-Ho-Viet/data/mai.RData")
saveRDS(mai, file = "D:/Viet/Thong-ke-Ho-Viet/data/mai.rds")
```

```{r}
library(leaflet)
nguyen_sff <- dat_sff %>% 
  mutate(prop.ho=(sum(subset(., Ho=="Nguyễn")[, "songuoi"])/sum(songuoi))*100) %>% 
  filter(Ho=="Nguyễn") %>% 
  st_as_sf(crs = 4326)

nguyen_sff$centroid <- sf::st_transform(nguyen_sff, 2154) %>% 
  st_centroid() %>% 
  # this is the crs from d, which has no EPSG code:
  st_transform(., '+proj=longlat +ellps=GRS80 +no_defs') %>%
  # since you want the centroids in a second geometry col:
  st_geometry()

nguyen <- nguyen_sff %>%
  mutate(x = map(centroid,1),
         y = map(centroid,2),
         x = as.numeric(x),
         y = as.numeric(y))

pal <- colorNumeric(palette = "viridis", reverse = TRUE, domain = nguyen_sff$songuoi_km, alpha = TRUE)

nguyen_sff %>%
    st_transform(crs = "+init=epsg:4326") %>%
    leaflet(width = "100%") %>%
    addProviderTiles(provider = "Esri.WorldStreetMap") %>%
    addPolygons(popup = ~ str_extract(NAME_3, "^([^,]*)"),
                stroke = FALSE,
                smoothFactor = 0,
                fillOpacity = 0.4,
                color = ~ pal(songuoi_km)) %>%
    addLegend("bottomright", 
              pal = pal, 
              values = ~ songuoi_km,
              title = "Mật độ người/km2",
              opacity = 1)

save(nguyen, file = "D:/Viet/Thong-ke-Ho-Viet/data/nguyen.RData")
saveRDS(nguyen, file = "D:/Viet/Thong-ke-Ho-Viet/data/nguyen.rds")
```

```{r}
library(leaflet)
tran_sff <- dat_sff %>% 
  mutate(prop.ho=(sum(subset(., Ho=="Trần")[, "songuoi"])/sum(songuoi))*100) %>% 
  filter(Ho=="Trần") %>% 
  st_as_sf(crs = 4326)

tran_sff$centroid <- sf::st_transform(tran_sff, 2154) %>% 
  st_centroid() %>% 
  # this is the crs from d, which has no EPSG code:
  st_transform(., '+proj=longlat +ellps=GRS80 +no_defs') %>%
  # since you want the centroids in a second geometry col:
  st_geometry()

tran <- tran_sff %>%
  mutate(x = map(centroid,1),
         y = map(centroid,2),
         x = as.numeric(x),
         y = as.numeric(y))

pal <- colorNumeric(palette = "viridis", reverse = TRUE, domain = tran_sff$songuoi_km, alpha = TRUE)

tran_sff %>%
    st_transform(crs = "+init=epsg:4326") %>%
    leaflet(width = "100%") %>%
    addProviderTiles(provider = "Esri.WorldStreetMap") %>%
    addPolygons(popup = ~ str_extract(NAME_3, "^([^,]*)"),
                stroke = FALSE,
                smoothFactor = 0,
                fillOpacity = 0.4,
                color = ~ pal(songuoi_km)) %>%
    addLegend("bottomright", 
              pal = pal, 
              values = ~ songuoi_km,
              title = "Mật độ người/km2",
              opacity = 1)

save(tran, file = "D:/Viet/Thong-ke-Ho-Viet/data/tran.RData")
saveRDS(tran, file = "D:/Viet/Thong-ke-Ho-Viet/data/tran.rds")
```


```{r}
library(sf)
library(leaflet)
# dat_sff <- dat_sff %>%
#   group_by(Ho) %>%
#   mutate(pro.ho=(sum(songuoi)/sum(danso))*100)
# 
# dat_sff <- dat_sff %>%
#   group_by(NAME_1, Ho) %>%
#   mutate(pro.tinh=(sum(songuoi)/sum(danso))*100)
# 
# dat_sff <- dat_sff %>%
#   group_by(NAME_1, NAME_2, Ho) %>%
#   mutate(pro.huyen=(sum(songuoi)/sum(danso))*100)

dat <- dat_sff %>% st_as_sf(crs = 4326)
# dat <- sf::st_simplify(dat)

# dat_sff$centroid <- sf::st_transform(dat_sff, 2154) %>% 
#   st_centroid() %>% 
#   # this is the crs from d, which has no EPSG code:
#   st_transform(., '+proj=longlat +ellps=GRS80 +no_defs') %>%
#   # since you want the centroids in a second geometry col:
#   st_geometry()
# 
# dat <- dat_sff %>%
#   mutate(x = map(centroid,1),
#          y = map(centroid,2),
#          x = as.numeric(x),
#          y = as.numeric(y))

ho <- data_ho_overall %>%
  arrange(desc(songuoi)) %>%
  mutate(pro.pop=(songuoi/sum(songuoi)*100))
ho <- head(ho, n=150)

ho10 <- head(ho, n=10)

dat <- dat %>% 
  filter(Ho %in% ho$Ho)
dat_spl <- rmapshaper::ms_simplify(dat, keep = 0.02)

dat1 <- dat %>% filter(Ho %in% c("Dương", "Lê", "Mai", "Ngô", "Nguyễn", "Phùng", "Trần"))

dat10 <- dat %>% 
  filter(Ho %in% ho10$Ho)

save(dat, file = "../../data/dat.RData")
saveRDS(dat, file = "../../data/dat.rds")
saveRDS(dat1, file = "../../data/dat1.rds")
saveRDS(dat10, file = "../../data/dat10.rds")
```
